Repairing cooperation through a cognitive intervention in the repeated Trust Game

### 2023-07-13 article.Rmd
plot improvements

```{r load-packages05, include = FALSE}
library(papaja)
library(kableExtra)
require(knitr)
#require(citr)
require(bookdown)

@@

Determining the other's goals, intentions, and decision-making process is fundamental to successful social interaction. This inference is however fraught with uncertainty, as we cannot directly observe these features, and may need extensive prior knowledge of the person. Absent a history of interaction, one may decide to trust the other's goals and intentions are aligned with one's own. The challenge of trust is that it is by construction a risky endeavour. If we deem a person trustworthy, we might take the risk of investing in the relationship, hoping for a collaborative outcome. If this trust is misplaced, it can come at a high cost. Not trusting others is also risky since opportunities for cooperation may be foregone.

@@

Evidence from the literature emphasises the importance of social trust in determining why some people fare better than others physically and mentally [@giordano_trust_2016 ; @meng_multilevel_2014]. It affects the health status of individuals by reinforcing social support networks, maintaining community norms and facilitating collective action. Research into the determinants of psychopathology has linked trust-based constructs to the emergence of mental health disorders. @fonagy_role_2014 identified epistemic trust, or the belief in the authenticity and personal relevance of interpersonally transmitted knowledge, as an important function of early attachment relationships. It allows the individual receiving social information to let go of their natural self-protective vigilance, which can become a pathological hypervigilance frequently observed after traumatic experiences and a key factor for the emergence of multiple mental health disorders [@fonagy_mentalizing_2017].

@@

A well-established paradigm in the study of trust is the repeated trust game [@joyce_trust_1995]. In this game, the "investor" decides how much of an endowment to send to the other player (the "trustee"). The amount that is sent is tripled and the trustee decides, in return, how much of the tripled amount to send back to the investor. To encourage the emergence and maintenance of trust in this setting, some studies focused on modifying the game mechanism, for example by introducing a third-party who monitors the actions of the other players [@charness_investment_2008; @fiedler_effect_2017]. Others chose to intervene directly on the participants. For example, @drazkowski_gratitude_2017 found that trust was increased when participants were asked to think about and write down five things that they were grateful for. @burnham_friend-or-foe_2000 found that trust was also increased when participants were primed with the concepts of friend and foe.

Whilst these interventions show that it is possible to improve cooperative outcomes at the start of the game, they do not address how to repair a breakdown of trust that might occur due to intentional or accidental non-cooperative actions by the players. Cooperative play in the repeated trust game can easily break down when there is a transgressive behavior, such as a nil or very low investment by the investor or a return of the trustee below the investment sent [@bendor_when_1991]. Such ruptures of cooperation appear frequently when the trustee suffers from social disorders such as Borderline Personality Disorder [BPD; @lieb_borderline_2004]. In theses situations, BPD trustees fail to engage in trust repairing behaviours such as coaxing the investor by signalling trustworthiness via sending high returns. This failure may be linked to the misperception that their low returns in the game are not violating social norms [@king-casas_rupture_2008].

In devising potential interventions to repair trust, we can derive inspiration from the cognitive interventions championed by successful psychological therapies that aim to improve aspects of social dysfunction in BPD patients. Whilst there is no proven pharmacological therapy for BPD, psychotherapies such as Mentalisation Based Therapy [@allen_handbook_2006-1] and Dialectical Behavior Therapy [@linehan_cognitive-behavioral_1993] have been shown to improve various dysfunctional behaviors in BPD patients, including those related to social interaction [@gunderson_borderline_2018]. However, response to these treatments is highly variable, and determining which interventions are effective for particular patients is challenging [@rudge_mechanisms_2020; @arch_longitudinal_2012]. One promising approach is the study of how specific components of psychotherapeutic treatment affect quantitative markers of behaviour such as those inferred through computational models [@huys_computational_2016; @reiter_neuro-cognitive_2021]. Combining the use of specific cognitive probes inspired by therapeutic interventions and computational models of behaviour may allow us to uncover the cognitive mechanisms targeted by common forms of psychotherapy. In turn, this may provide the basis for choosing effective psychotherapeutic interventions for given individuals.

In this study, we assess the effectiveness of a cognitive intervention aimed at repairing the potential breakdown of cooperation from a one-off low investment sent by an computerised investor. The intervention focuses on explaining the potential harm from reciprocating non-cooperative actions and suggesting a non-impulsive course of action to coax the investor back into cooperation. Participants are given the role of the trustee and are randomly assigned to either a control or intervention group. They play two instances of the repeated trust game with two different investors. After the first instance of the game, they either receive a cognitive intervention (intervention condition) or perform an unrelated task solving anagrams (control condition). In reality, participants face the same computerised agent in both instances, which is programmed to play according to a hidden Markov model fitted to real players' data. We explore whether the intervention has an effect on the behaviour of the human trustee and whether this effect transfers to a different game (repeated Prisoner's Dilemma) and facing a seemingly new player.

@@

A total of 318 participants were recruited on the Prolific Academic platform (prolific.co). The mean age of participants was `r round(mean(as.numeric(d_finished$exitSurvey.age), na.rm=TRUE),1)` years. Participants were paid a fixed fee of £5 plus a bonus payment dependent on their performance.<!--MS: state average bonus--> The experiment had a 2 (Condition: Intervention or Control) by 3 (Game: Trust-Game Pre Intervention, Trust-Game Post Intervention, Prisoner's Dilemma Post Intervention) design, with repeated measures on the second factor. Participants were randomly assigned to one of the two levels of the first factor.

## Tasks and Measures

### Repeated Trust Game

Participants played a 15-round repeated trust game [@joyce_trust_1995] in the trustee role against a computer-programmed investor. On each round the investor was endowed with 20 units and decided how much of that endowment to invest. This investment is tripled and the trustee then decides how to split this amount between them and the investor. If the trustee returns more than one third of the amount, the investor makes a gain. The Nash equilibrium for a single-round version is for the investor to send nothing. In the repeated version, rewards for both players are maximised if they build trust and share the benefits of the the investment multiplied by three. An investor who has been rewarded for taking the risk of sending an investment is more likely to invest more on future rounds. An investor obtaining a low return on their investment may choose to reduce future investment and thereby reduce both players’ gains.

@@

The strategy of the computerised investor was modelled on behaviour of human investors in a 10-round RTG.<!-- reference to data -->.  Using this data, we estimated a hidden Markov model (HMM) on investors' behaviour with three latent states, reflecting "low-trust", "medium-trust", or "high-trust". Each latent state was associated with a distribution over all possible investor actions (from 0 to 20 investment) that reflected the amount of trust (Figure \ref{fig:HMMPanels}.A). Over rounds, the investor can move between states, and the probability of these transitions was modelled as a function of the net return in the previous round (see Figure \ref{fig:HMMPanels}.B).

In order to probe efforts to repair trust, the computerised agent was programmed to provide a low return on round 12 pre-intervention or 13 post-intervention. On all other rounds, the agent's actions were derived from the HMM (disregarding the participant's response on round 12).

### Repeated Prisoner's Dillema

Participants played 7 rounds of a repeated Prisoner's Dilemma. In each round, participants could choose a cooperative action with a a reward of 5 if the player also cooperated and the a reward of 1 if not, or a non-cooperative action that would yield a 7 points if the other person chooses the cooperative action and 2 points if not. The Nash equilibrium for a single-round version is to choose the non-cooperative action.

The computerised agent was programmed to act according to a tit-For-tat strategy [@axelrod_evolution_1981], starting with a cooperative action and then mirroring what the other player chose in the preceding round. On round 4, the agent was pre-programmed to choose the defect action, regardless of the participant's preceding action.

## Intervention

```{r timeline, fig.cap = "Experiment overview. After playing 15 rounds of a repeated trust game (RTG) as trustee, participants were randomized to either receive the control or active intervention. They then played the second set of 15 rounds of RTG (again as trustee) to examine intervention effects, and 7 rounds of a repeated Prisoner’s Dilemma to examine generalization of intervention effects. Finally, participants answered questionnaires and were debriefed.", fig.align='center', out.width="80%"}

At the start of the experiment (Figure \ref{fig:timeline}), participants provided informed consent and were instructed the study would consist of three phases in which they would face a different other player. Participants were told their goal was to maximise the number of points in all phases. They were not told the number of rounds of each phase. Phase one was a 15 round repeated Trust Game (RTG) in which participants took the role of trustee, facing the same investor over all 15 rounds. On each round, after being informed about the amount sent by the investor, participants were asked to provide feedback, with participants in the Intervention condition rating their feeling in terms of valence (from negative to positive) and arousal (from low to high), and participants in the Control condition rating the investment in terms of speed (from slow to fast) and magnitude (from low to high). These ratings were made by indicating a point on a two-dimensional grid. Participants then decided on how much of the tripled investment to return to the investor, before continuing to the next round. After completing 15 rounds of the RTG, participants rated how cooperative, selfish, trustworthy and friendly they perceived their interaction partner (all on a scale from 1 to 10). After phase one, participants in the intervention condition completed the intervention, and participants in the control condition solved anagrams. Subsequent phase two was similar to phase one, with participants being told they would face a new player. Phase three consisted of 7 rounds of the repeated prisoner's dilemma game (RPD), with participants informed they would face a third player. Participants then completed questionnaires related to mentalising abilities (RFQ8), emotion regulation (DERS), and BPD traits (PAI-BOR).<!-- MS: need to include references for these scales --> They were then asked about the strategy in the games, as well as whether they thought the other players were human or computer agents. They were then debriefed and thanked for their participation.

## Statistical analysis

@@

To model participants' returns in the RTG across games and conditions, we fit various hidden Markov models to participants' returns using the depmixS4 package [@visser_depmixs4_2021] for R. The transition between latent states is assumed to depend on the investment received and a dummy variable to characterise the group that the participant belongs to. Details on how the models are constructed can be found in the supplement. We fit models with different numbers of hidden states, and use the Bayesian Information Criterion [@schwarz_estimating_1978-1] to select the best model.

# Behavioural results

@@

Average investments and returns prior to the "defection round" (Figure \ref{fig:gamesPlot}) were within the range of reported investments (40-60% of endowment) and returns (35-50% of total yield) in the literature [@charness_investment_2008; @fiedler_social_2011].

<!--
Estimated marginal means of percentage returns from fitting the mixed effects model specified in the methods section, as well as the distribution of the returns for participants across all trials are presented in Figure \ref{fig:violinPanels}.A.-->

@@

ggplot(avg_ret_df ,aes(x=as.factor(roundNum),group=gameNum.f, color = gameNum.f)) +
  stat_summary(fun = "mean", geom = "line", aes(y=ret_pct)) +
  stat_summary(fun.data = "mean_se", geom = "errorbar", alpha =0.3, width = 0.2, aes(y=ret_pct) )+
  labs(x = "Round",
       y = "Percentage Return",
       color = "Game") +
  # ylab("Percentage Return") +
  # xlab("Round") +
  theme_bw() +
  theme(legend.position = "bottom") +
  facet_wrap(~condition.f)

@@

mod_returns_pct <- mixed( ret_pct ~ gameNum.f*condition.f*inv_scaled+ (1 + gameNum.f| id), avg_ret_df, REML= TRUE, method="KR")
anova(mod_returns_pct)
summary(mod_returns_pct)

@@

<!-- MS: double check statistics; I get different df when using KR in mixed -->
Mixed-effects analysis shows a significant main effect of Condition (intervention vs. control), `r papaja::apa_print(mod_returns_pct)$full_result$condition_f` <!--$F(1,315.31) = 9.52$, $p = .002$-->, with higher percentage returns in the intervention compared to the control condition. Importantly, we also find an interaction between Condition and Game-number (pre- vs. post-intervention), `r papaja::apa_print(mod_returns_pct)$full_result$gameNum_f_condition_f` <!--$F(1,314.2) = 26.9$, $p < .001$-->. Post-hoc tests show an increase in the percentage returned in the intervention condition, pre - post `r papaja::apa_print(prepost_bycond)$full_result$Intervention_Pre_post`, but a decrease in the control condition, `r papaja::apa_print(prepost_bycond)$full_result$Control_Pre_post` (see Figure \ref{fig:violinPanels}.A). This indicates the intervention was effective in increasing cooperative behaviour. There was also a significant main effect of Investment, `r papaja::apa_print(mod_returns_pct)$full_result$inv_scaled` <!--$F(1,9208) = 373.6$, $p < 0.001$-->, such that higher investments were associated with higher percentage returns. An Investment by Condition interaction, `r papaja::apa_print(mod_returns_pct)$full_result$condition_f_inv_scaled` <!--$F(1,9207) = 45.38$, $p < .001$-->, indicates the positive effect of investment on percentage returns was greater in the control than intervention condition. <!-- This suggests that the higher returns we see in the intervention group are not simply due to higher investments.--> There was also an Investment by Game-number interaction, `r papaja::apa_print(mod_returns_pct)$full_result$gameNum_f_inv_scaled`. Finally, we find a three way interaction between Game-number, Condition and Investment, `r papaja::apa_print(mod_returns_pct)$full_result$gameNum_f_condition_f_inv_scaled`, <!--($F(1,8988)= 24.6, p < 0.001$),--> showing that the differentiated effect of the investment on the proportion returned by condition is itself moderated by the Game-number.

<!--
group had higher percentage returns than the control group.
We find an interaction effect between Condition (intervention vs. control) and Game-number (pre- vs. post-manipulation) ($F(1,314.2) = 26.9, p < 0.001$). Post-hoc tests show that there was an increase in the percentage returned in the intervention group (`r papaja::apa_print(prepost_bycond)$statistic$Intervention_Pre_post`) but not in the control group (`r papaja::apa_print(prepost_bycond)$statistic$Control_Pre_post`). We also found a significant main effect for Condition ( $F(1,315.31) = 9.52, p = 0.002)$), such that the intervention group had higher percentage returns than the control group. There was also a significant main effect of Investment( $F(1,9208) = 373.6, p < 0.001$) such that higher investments were associated with higher percentage returns.

@@

-->

```{r, include=FALSE, cache=TRUE}
mod_invs <- mixed( investment ~ gameNum.f*condition.f  + (1 + gameNum.f| id), avg_ret_df, method="KR")
summary(mod_invs)
anova(mod_invs)

#cond_effect_inv <- pairs(emmeans::emmeans(mod_invs, c("condition.f"), by = "gameNum.f", pbkrtest.limit = 9570))

@@

We next analysed HMM agent investments via a linear-mixed effects model with fixed effects for Condition (intervention or control), Game-number (pre or post intervention), as well as interaction between Condition and Game-number, and participant-wise random intercepts and random slopes for Game-number. This shows a main effect of Condition, `r papaja::apa_print(mod_invs)$full_result$condition_f`<!--$F(1,317) = 8.7$, $p = .003$-->, and Game-number, `r papaja::apa_print(mod_invs)$full_result$gameNum_f`<!--$F(1,317) = 8.3$, $p = .004$-->. As can be seen in Figure \ref{fig:violinPanels}.B, investment was higher in the intervention compared to the control condition across games, and higher in the second game compared to first across conditions.

Taken together, we find that participants in the control condition sent *lower* percentage returns in the second game, despite the HMM investor sending on average higher investments in the second game. Those in the intervention group returned *higher* percentage returns in the second game, with the investor also sending higher investments. These higher returns in the intervention group compared to the control group were not purely driven by reciprocity towards higher investments, since we find a Condition by Game-number interaction whilst controlling for investment in the model, and a reduced effect of Investment in the Intervention condition. <!-- since we find this interaction effect whilst controlling for investments in our model of participants returns.-->

<!-- ## Looking at pre- and post-defection trials only -->

@@

mod_returns_pre <- mixed( ret_pct ~ gameNum.f*condition.f*inv_scaled+ (1 + gameNum.f| id), pre_int_data, REML= TRUE, method="KR")
anova(mod_returns_pre)
summary(mod_returns_pre)

@@

We next analysed returns separately for rounds prior to the pre-programmed defection (rounds 1 to 11 pre-intervention and 1 to 12 post-intervention) and rounds after (rounds 12 to 15 pre-intervention and rounds 13 to 15 post-intervention). Applying the same mixed effects model as before to returns before defection, largely replicates results over all trials. Full results are provided in the supplement. Participants in the intervention condition increased their returns in the second game, `r papaja::apa_print(pairs_pre_int)$full_result$Intervention_Pre_post`. There was no evidence that participants in the control condition changed their returns in pre-defection trials between the two phases of the RTG, `r papaja::apa_print(pairs_pre_int)$full_result$Control_Pre_post`.

<!--
investments Analysis  and fit the same linear mixed effects model to percentage returns as the one for the full data. Estimated marginal means of percentage returns as well as the distribution of the returns for participants across all trials are presented in Figure \ref{fig:violinPanels}.B. We do find an interaction effect between Condition and Game number ($F(1,317.99) = 17.1, p < 0.001$). This was due to an increase in the percentage returned pre low-investment trial in the intervention condition (`r apa_print(pairs_pre_int)$full_result$Intervention_Pre_post`) but not in the control group (`r apa_print(pairs_pre_int)$full_result$Control_Pre_post`). This indicates that participants were sending back higher returns in the intervention condition but not in the control condition even before the defection trial.
-->

@@

@@

```{r violinPanels, fig.cap="A: Marginal means and distributions of percentage trustee returns over all rounds, shown across participants by Game number and Condition. B: Marginal means and distributions of investments over all rounds for HMM investors, by Game number and Condition. C: Marginal means and distributions of percentage trustee returns across all participants for pre-defection trials only, by Game number and Condition. D: Marginal means and distributions of percentage trustee returns across all participants for post-defection trials only, by Game number and Condition.", fig.align="center", fig.width=6, fig.height = 6}

ggpubr::ggarrange(pctRet_all, inv_all, pctRet_pre, pctRet_post, nrow=2, ncol=2, labels = c("A", "B", "C", "D"), common.legend = TRUE, legend = "bottom")
```

The intervention focused on repairing cooperation after a "transgressive action" by the investor. To assess whether the intervention succeeded, we next considered the returns by participants after the pre-programmed transgression (Figure \ref{fig:violinPanels}.D). <!-- We fit the same linear mixed effects model to percentage returns as the one for the full data while restricting the data to rounds 12 to 15 in the first game, and 13 to 15 in the second game. These trials represent all trials after the pre-programmed defection of the investor (low investment of $2$) until the end of the game (round 15).
-->
<!-- Figure \ref{fig:violinPanels}.D shows the marginal means of trustee returns and their distribution by Game-number and Condition.--> A mixed-model analysis shows similar results to those of the model applied to all returns (see SI for full results). The was again a significant interaction between Condition and Game-number and Condition. Participants in the control condition decreased their post-transgression returns from the first to the second phase of the RTG, `r papaja::apa_print(pairs(postcoax_interaction, by = "condition.f", reverse=FALSE))$full_result$Control_Pre_post`. There was no change for participants in the intervention condition, `r papaja::apa_print(pairs(postcoax_interaction, by = "condition.f", reverse=FALSE))$full_result$Intervention_Pre_post`. So, whilst the intervention did not appear to move participants to increase their returns after a transgression by the investor, it may have countered the decrease in returns showed by participants in the control condition.

@@

# p <- datCoax %>%
#   ggplot(aes(x=roundNum)) + stat_summary(fun = "mean", geom = "line", aes(y=emo_scaled_x, color = "Valence")) + stat_summary(fun = "mean", geom = "line", aes(y=emo_scaled_y, color = "Arousal")) + stat_summary(fun = "mean", geom = "line", aes(y=inv_scaled, color = "Investment")) + facet_wrap(~gameNum_f, scales = "free_x")

>>>>

```{r load-packages05, include = FALSE}
library(papaja)
library(kableExtra)
require(knitr)
require(citr)
require(bookdown)

@@

At the core of social interaction is figuring out the goals, intentions, and decision-making process of the interaction partner. This inference is however fraught with uncertainty, as we cannot reasonably observe these features without prior knowledge of the person. Absent a history of interaction, one needs to decide whether to take the risk of trusting others. The challenge of trust is that it is by construction a risky endeavour. For example, if we deem a person trustworthy, we might decide to take the risk of investing in the relationship hoping for a collaborative outcome. If we misplace our trust, this can come at a high cost to us. Not trusting others is also risky since opportunities for cooperation may be foregone.

@@

Evidence from the literature emphasises the importance of social trust in determining why some people fare better than others physically and mentally [@giordano_trust_2016 ; @meng_multilevel_2014]. It affects the health status of individuals through reinforcing social support networks, maintaining community norms and facilitating collective action. Research into the determinants of psychopathology has linked trust-based constructs to the emergence of mental health disorders. @fonagy_role_2014 identified epistemic trust, or the belief in the authenticity and personal relevance of interpersonally transmitted knowledge, as an important function of early attachment relationships. It allows the individual receiving social information to let go of their natural self-protective vigilance, which can become a pathological hypervigilance frequently observed after traumatic experiences and a key factor for the emergence of multiple mental health disorders [@fonagy_mentalizing_2017].

@@

A well-established paradigm in the study of trust is the repeated trust game [@joyce_trust_1995]. To encourage the emergence and maintenance of trust in this setting, some studies focused on modifying the game mechanism [@charness_investment_2008; @fiedler_effect_2017]. Others chose to intervene directly on the participants. For example, @drazkowski_gratitude_2017 found that trust was increased when participants were asked to think about and write down five things that they were grateful for. @burnham_friend-or-foe_2000 found that trust was also increased when participants were primed with the concepts of friend and foe.

Whilst these interventions show that it is possible to improve cooperative outcomes at the start of the game, they did not address how to repair a breakdown of trust that might occur due to intentional or accidental non-cooperative actions by the players. Indeed, cooperative play in the repeated trust game can easily break down when there is a transgressive behavior, such as a nil or very low investment by the investor after cooperation has been established, or a return of the trustee below the investment sent. Such ruptures of cooperation appear frequently when the trustee suffers from mental health disorders affecting the social domain such as Borderline Personality Disorder [@lieb_borderline_2004]. In theses situations, BPD trustees fail to engage in trust repairing behavior such as coaxing the investor through sending high return to signal trustworthiness, and this failure may be linked to a failure to perceive their low returns in the game as a violation of social norms [@king-casas_rupture_2008].

In devising potential interventions to repair trust, we can derive inspiration from the cognitive interventions championed by successful psychological therapies that aim to improve aspects of interpersonal dysfunction in BPD patients. Whilst there is no proven pharmacological therapy for BPD, some forms of psychotherapy such as Mentalisation Based Therapy [@allen_handbook_2006-1] and Dialectical Behavior Therapy [@linehan_cognitive-behavioral_1993], have been clinically validated as efficacious approaches to improve various dysfunctional behaviors in BPD patients, including those related to social interaction [@gunderson_borderline_2018]. However, these therapies suffer from a high variability in treatment responses. Determining which interventions are effective for particular patients has been very challenging [@rudge_mechanisms_2020; @arch_longitudinal_2012]. One promising approach is the study of how specific psychotherapeutic treatment components affect quantitative markers of behavior such as those inferred through computational modelling techniques [@huys_computational_2016; @reiter_neuro-cognitive_2021]. As such, using specific cognitive probes inspired by therapeutic interventions complemented by computational modelling of behavior in tasks, may allow us to uncover cognitive mechanisms targeted by common forms of psychotherapy. This may in turn form the basis for targeting of existing psychotherapies to different individuals.

In this study, we assess the effectiveness of a cognitive intervention aimed at repairing the potential breakdown of cooperation from a pre-programmed, one-off low investment sent by an computer investor. The intervention focuses on explaining the potential harm from reciprocating non-cooperative actions and suggesting a non-impulsive course of action to coax the investor back into cooperation. Participants are given the role of the trustee and are randomly assigned to either a control group or an intervention group. They play two instances of the repeated Trust Game facing what they believe are two different players. In between games, they either receive a cognitive intervention (intervention condition) or not (control condition, where participants are asked to solve anagrams). In reality, participants face the same computerised agent in both instances, which is programmed to play according to an HMM fitted to real players data. We explore whether the intervention has an effect on the behaviour of the trustee and whether the learning is transferred to a new repeated Prisoner's Dilemma game when facing a seemingly new player.

@@

A total of 318 participants were recruited on the Prolific Academic platform (prolific.co). The mean age of participants was `r round(mean(as.numeric(d_finished$exitSurvey.age), na.rm=TRUE),1)` years. Participants were paid a fixed fee of £5 plus a bonus dependent on their performance. The experiment had a 2 (Condition: Intervention or Control) by 2 (Game : Pre or Post Intervention) design, with repeated measures on the first factor. Participants were randomly assigned to one of the two levels of the second factor.

```{r timeline, fig.cap = "A) Experiment overview. After playing 15 rounds of a repeated trust game (RTG) as trustee, participants were randomized to either receive the control or active intervention. They then played the second set of 15 rounds of RTG (again as trustee) to examine intervention effects, and 7 rounds of a repeated prisoner’s dilemma to examine generalization of intervention effects. Finally, participants answered questionnaires and were debriefed.", fig.align='center', out.width="80%"}

The experiment as shown in Figure \ref{fig:timeline} began with participants briefed there will be three phases, and that they would face a new opponent in each phase. Phase one was a 15 round Trust Game (RTG) in which participants had the role of the trustee, facing the same investor opponent. Participants were then assigned to either the control or intervention conditions. Phase two was similar to Phase one, with participants being told they faced a new player. In both phases, participants were asked for feedback in each round after seeing the investment. The nature of the feedback solicited varied by condition, either emotion-related (intervention) or unrelated (control). Participants rated their interaction partner on various attributes after each phase. More details on these attributes, nature of feedback and exact instructions are found in the supplement. Phase three involved 7 rounds of the repeated prisoner's dilemma game (RPD) with a third player. Participants were explicitly told to aim to maximise the number of points in all phases. The total number of rounds was not communicated to the participants in any of the games played.

Post-experiment, participants completed questionnaires related to mentalising abilities, emotion regulation and BPD traits (see details in supplement). They were then quizzed on their game-play, including whether they detected that all opponents were computer agents. The reveal explained that they were facing identical AI agents and that any perceived opponent changes were due to participants' own behavior.

## Interventions

## Design of the AI agents in the RTG and RPD

@@

```
The role of the investor was in reality played by an adaptive artificial agent whose strategy was modelled on the behaviour of human participants taking the role of the investor in the RTG. Using data from human dyadic interaction in the 10 round trust game, we estimate a Hidden Markov Model (HMM) to characterise the investor's behavior as emanating from a small number of latent states [@rabiner_hmm_1989]. In this instance, for both iterations of the repeated trust game, the HMM used was identical: it had three states that can be described as "low-trust", "medium-trust" and "high-trust". Each latent state was associated with a distribution over all possible investor actions (from 0 to 20 investment) that reflected the amount of trust, as presented in Figure \ref{fig:HMMPanels}.A. Transition probabilities between these latent states were modelled as a function of the net return in the previous round conditional on the investor's state as shown in Figure \ref{fig:HMMPanels}.B.

<!-- In order to probe efforts to repair trust, we introduced rounds where the investor deviated from the policy provided by the HMM, and chose to send a low investment in the Trust Game, or chose the non-cooperative action in the Prisoner's Dilemma. In the trust games, we programmed the HMM agent to send a very low investment (2 out of 20) to the trustee on specific rounds (round 12 pre-intervention and 13 post-intervention). For the repeated Prisoner's Dilemma game, we programmed the agent to choose the non-cooperative action on round 4. After each "defection" round, the agents were programmed to proceed their policy from before that round, effectively ignoring the participants' response to their defection. For instance, the HMM investor would send an investment in the post defection round that was consistent with the outcome of the round immediately prior to that round. For the RPD, the agent would continue to play tit-for-tat, but also ignoring what happened in the defection round, and basing its action on mirroring what the other player did in the round immediately prior to the defection round. This ignorance of reactions to the defection rounds was chosen to reflect an accidental defection, with the Investor being open to repair subsequent interaction.  -->

In order to probe efforts to repair trust, we simulated trust breaches in both Trust Games and the Prisoner's Dilemma. The AI was programmed to break trust on specific rounds (12 pre-intervention, 13 post-intervention in the Trust Game; round 4 in Prisoner's Dilemma). After these "defection" rounds, the AI returned to its previous strategy, disregarding the participant's response to the breach. This simulated an accidental defection and a readiness to repair the interaction, providing insights into trust repair dynamics.

<!-- ## Computational analysis of trustee returns -->

<!-- In order to characterise the effect of the intervention on the trustees strategy, we fit a Hidden Markov Model to the percentage returns of the trustees for the RTG data. The model allows for switching between latent trustee states over time. Hidden Markov models assume that an observable trustee return at time $t$ depends on a latent trustee state at time $t$. It is assumed that the latent state at time $t$ depends on the latent state at the previous time $t-1$. The model is specified by the state-conditional action distributions (these are provided by the likelihood of the fitted models), an initial state distribution (the distribution over the strategies at the initial round), and the state-transition probabilities (probability of switching from one state/strategy to another). Initial state probabilities and the transition probabilities were estimated with the depmixS4 package [@visser_depmixs4_2021]. -->

## Analysis of participants returns in the RTG

@@

To model participants returns in the RTG across games and conditions, we fit various hidden Markov models to participants returns using the depmixS4 package [@visser_depmixs4_2021]. The transition between latent states is assumed to depend on the investment received and a dummy variable to characterise the group that the participant belongs to. Details on how the models are constructed as well as the various contrast codes used for the dummy variables are found in the supplement. We fit models with different numbers of hidden states, and use the Bayesian Information Criterion [@schwarz_estimating_1978-1] to select the best model.

# Behavioral results

@@

Average investments and returns prior to the defection round (Figure \ref{fig:gamesPlot}) were within the range of reported investments (40-60% of endowment) and returns (35-50% of total yield) in the literature [@charness_investment_2008; @fiedler_social_2011]. Estimated marginal means of percentage returns from fitting the mixed effects model specified in the methods section, as well as the distribution of the returns for participants across all trials are presented in Figure \ref{fig:violinPanels}.A.

@@

ggplot(avg_ret_df, aes(x=as.factor(roundNum), y=ret_pct, group=gameNum.f, color = gameNum.f)) +
  stat_summary(fun = mean, geom = "line") +
  stat_summary(fun.data = mean_se, geom = "ribbon", aes(ymin=..ymin.., ymax=..ymax..), 
               alpha = 0.3, fill = "grey", linetype = 0) +
  labs(x = "Round",
       y = "Percentage Return",
       color = "Game") +
  theme_bw() +
  theme(legend.position = "bottom") +
  facet_wrap(~condition.f) +
  scale_color_manual(values = c("darkblue", "green"))

<!-- Plotting individual differences   -->
```{r, include =F}
ret_post_int <- avg_ret_df %>% filter(roundNum == 13 , gameNum.f == "post")

# plot(jitter(ret_post_int$returns), main="Jitter Plot", ylab="Returns", pch=19, col="blue")
# plot(ecdf(ret_post_int$returns), main="ECDF", xlab="Returns", ylab="Cumulative Probability")
# 
# ggplot(ret_post_int , aes(x = "", y = returns)) +
#   geom_violin(fill="lightblue") +
#   geom_boxplot(width=0.1, fill="white") +
#   theme_minimal() +
#   labs(x = NULL, y = "Returns", title = "Violin plot")

library(ggplot2)
ggplot(ret_post_int, aes(x = "", y = ret_pct)) +
  geom_violin(fill="lightblue") +
  geom_jitter(width = 0.2, height = 0.02, size = 1, col = "black") +
  geom_hline(yintercept = 0.33, color = "red", linetype = "dashed") +
  theme_bw() +
  facet_wrap(~condition.f) +
  labs(x = NULL, y = "% Return", title = "")

@@

mod_returns_pct <- mixed( ret_pct ~ gameNum.f*condition.f*inv_scaled+ (1 + gameNum.f| id), avg_ret_df, REML= TRUE)
anova(mod_returns_pct)
summary(mod_returns_pct)

@@

We find an interaction effect between Condition (intervention vs. control) and Game-number (pre- vs. post-manipulation) ($F(1,314.2) = 26.9, p < 0.001$). Post-hoc tests show that there was an increase in the percentage returned in the intervention group (`r papaja::apa_print(prepost_bycond)$statistic$Intervention_Pre_post`) but not in the control group (`r papaja::apa_print(prepost_bycond)$statistic$Control_Pre_post`). We also found a significant main effect for Condition ( $F(1,315.31) = 9.52, p = 0.002)$), such that the intervention group had higher percentage returns than the control group. There was also a significant main effect of Investment( $F(1,9208) = 373.6, p < 0.001$) such that higher investments were associated with higher percentage returns.

@@

```{r, include=FALSE, cache=TRUE}
mod_invs <- mixed( investment ~ gameNum.f*condition.f  + (1 + gameNum.f| id), avg_ret_df )
summary(mod_invs)
anova(mod_invs)

@@

The HMM agent exhibited a differentiated behavior between games and conditions. When examining the HMM agent investments, we fit a linear-mixed effects model, with fixed effects for Condition (intervention or control), Game-number (pre or post intervention), as well as interaction between Condition and Game-number, and participant-wise random intercepts and random slopes for Game-number. We find a main effect for both Condition ($F(1,317) = 8.7, p = 0.003$) and Game-number($F(1,317) = 8.3, p = 0.004$). As can be seen in Figure \ref{fig:violinPanels}.D, investment was higher in the intervention compared to the control condition across games ($p=0.003$). Across conditions, investment was also higher in the second game compared to the first ($p=0.003$).

In summary, participants in the control group sent back **lower** percentage returns in the second game (`r apa_print(prepost_bycond)$full_result$Control_Pre_post`) despite the HMM investor sending, on average higher investments. Those in the intervention group returned **higher** percentage returns in the second game (`r apa_print(prepost_bycond)$full_result$Intervention_Pre_post`), with the investor also sending higher investments. These higher returns in the intervention group compared to the control group were not purely driven by reciprocity towards higher investments since we find this interaction effect whilst controlling for investments in our model of participants returns.

## Looking at pre- and post-defection trials only

@@

mod_returns_pre <- mixed( ret_pct ~ gameNum.f*condition.f*inv_scaled+ (1 + gameNum.f| id), pre_int_data, REML= TRUE)
anova(mod_returns_pre)
summary(mod_returns_pre)

@@

We restrict our analysis to the rounds prior to the pre-programmed defection (rounds 1 to 11 pre-manipulation and 1 to 12 post-manipulation) and fit the same linear mixed effects model to percentage returns as the one for the full data. Estimated marginal means of percentage returns as well as the distribution of the returns for participants across all trials are presented in Figure \ref{fig:violinPanels}.B. We do find an interaction effect between Condition and Game number ($F(1,317.99) = 17.1, p < 0.001$). This was due to an increase in the percentage returned pre low-investment trial in the intervention condition (`r apa_print(pairs_pre_int)$full_result$Intervention_Pre_post`) but not in the control group (`r apa_print(pairs_pre_int)$full_result$Control_Pre_post`). This indicates that participants were sending back higher returns in the intervention condition but not in the control condition even before the defection trial.

@@

The interventions focused on repairing cooperation in the wake of a transgressive action by the investor. To assess the effect of the intervention, we can therefore consider the percentage return after the "transgression" trials. We fit the same linear mixed effects model to percentage returns as the one for the full data while restricting the data to rounds 12 to 15 in the first game, and 13 to 15 in the second game. These trials represent all trials after the pre-programmed defection of the investor (low investment of $2$) until the end of the game (round 15).

@@

```{r violinPanels, fig.cap="A: Marginal means and distributions of percentage trustee returns over all rounds, shown across participants by Game number and Condition. B: Marginal means and distributions of percentage trustee returns across all participants for pre-defection trials only, by Game number and Condition. C: Marginal means and distributions of percentage trustee returns across all participants for post-defection trials only, by Game number and Condition. D: Marginal means and distributions of investments over all rounds for HMM agents playing the role of the investor, by Game number and Condition", fig.align="center", fig.width=6, fig.height = 6}

ggpubr::ggarrange(pctRet_all, pctRet_pre, pctRet_post, inv_all, nrow=2, ncol=2, labels = c("A", "B", "C", "D"), common.legend = TRUE, legend = "bottom")
```

Figure \ref{fig:violinPanels}.C shows the marginal means of trustee returns and their distribution by Game-number and Condition. There was a significant interaction between Game-number and Condition. Participants sent back lower returns between games in the control condition (`r papaja::apa_print(pairs(postcoax_interaction, by = "condition.f", reverse=FALSE))$full_result$Control_Pre_post`), but no such difference in the intervention condition (`r papaja::apa_print(pairs(postcoax_interaction, by = "condition.f", reverse=FALSE))$full_result$Intervention_Pre_post`).

@@
